\documentclass{report}
\usepackage{fullpage}

\usepackage{physics}
\usepackage{tikz}
\usepackage{amsmath, graphicx, xspace}
\usepackage{amssymb, amsthm}
\usepackage{enumitem}
\usepackage{siunitx}
\usepackage{hyperref}

% \usepackage[backend=bibtex,style=verbose-trad1, citestyle=numeric]{biblatex}


\newcommand\xp{x\prime\xspace}
\newcommand\yp{y\prime\xspace}
\newcommand\vp{v\prime\xspace}

\newcommand{\bb}[1]{\mathbf{#1}}



\title{Orbits 5: A Graphical N-body Gravitational Simulator}
\author{Leonardo Carnovale}

\begin{document}
  \maketitle

  \tableofcontents
  \chapter{Requirements and Use Cases}
  \section{Requirements}
      An informal depiction of the requirements of the software.
      \subsection{Functional Requirements}
        The software must be able to:
        \begin{itemize}
          \item Represent a system of particles by an array of
            N-dimensional positions, and optionally velocity, mass,
            radius and any other desired quantity, each as an array of
            scalars or arrays.
          \item Perform a simulation by executing an arbitrary step
            function.
          \item Perform simulations on any system of particles.
          \item Perform simulations with an arbitray force.
            between particles, and arbitrary collision behaviour.
          \item Model collisions between particles where the radii
            of spherical particles intersect.
          \item 
            Create buffers that, for every frame, can replicate a system to
            any desired degree of detail. 
          \item 
            Perform a simulation in real time (or as close as possible), or load frames
            into memory as a buffer, which can be recalled later.
          \item 
            Render a system or frame into screen coordinates 
            (x and y position, angle, radius etc.)
          \item 
            Draw a render of a system or frame onto a GUI interface, that supports mouse and keyboard
            interaction.
          \item 
            User defined modules for physics functions, rendering, and drawing should be able to
            made seperately to the main modules, that can be used instead of default modules.
        \end{itemize}
        The various separable modules should be as independent as possible. Apart from importing
        functions to be used by upper-level modules, lower level modules should not rely
        on any specific attributes of other modules. 

      \subsection{Non-Functional Requirements}
        The primary constraint and goal of the simulation is adequate speed and efficiency of calculations.
        It is likely achieving a performance better than order $O(n^2)$ will be impossible for a true
        N-body simulation. Hence, other types of simulation methods should be available aswell.

        Accepted input (ie a given system) does not have to be expected to be valid for
        every simulation, but the only modules that should have any strict requirements
        are the Physics, Camera, and Drawing modules. These may require certain attribtues of
        systems such as mass, radius, colour etc.
        % \begin{itemize}
        %   \item 
        % \end{itemize}
      \subsection{Analysis of Requirements}
        The described constraints, aswell as decisions made prior to defining the requirements,
        force the simulator to be spread across multiple modules and files. Modules regarding
        core algorithms like performing steps, representing steps, and representing buffers,
        should be encapsulated in functions and methods for use in other potentially
        arbitrary user defined modules. 
        
        The user defined modules will need to allow
        for any \textit{type} of simulation at the very least. In addition, if wanted
        the user should be able to define their own modules to render and display simulations.
        The rendering and drawing methods to be included with the simulation should be
        comprehensive and adequately explore the range of any possible user-defined methods.
      \subsection{User Stories}
        
  \chapter{Simulator and Graphics API}
    \section{Relationship between simulator and API}
      The graphical display should not be a design specific to the
      operation of the simulator, so as to allow significant improvement
      and modification in future versions. This is to say that the
      simulator and the graphics API should operate independently, and
      communication or interaction should mainly be from the simulator to the camera,
      then to the API. The simulator will need to prioritise simulation, and spend all
      its time on this task. It should not try to optimize output for the API,
      this can be done within the API.

    \section{Function of the Simulator}
      The job of the simulator is simply to calculate the position of all objects
      after a time step interval. The main program will be frequently interacting with it.
      The simulator should have the following methods:
      \begin{itemize}
        \item Add an object to the simulation.
        \item Remove an object from the simulation.
        \item Change the default time step interval.
        \item Perform a step.
        \item Perform $n$ steps.
      \end{itemize}
      There are multiple ways to perform this simulation, and the current preference
      is the leapfrog method. Other methods should be available to be chosen at creation of
      the simulator object.

      The above methods will likely not need to have a return value. All information about the
      simulator such as the current time step, number of particles, or the information about a
      particular particle, should be available via additional `get' methods.
    \section{Function of the API}
      The API must have the following functionality as a minimum:
      \begin{itemize}
        \item Display a coloured pixel at a given point.
        \item Display a coloured circle or ellipse at a given location, filled or unfilled.
        \item Display a coloured regular or irregular polygon at a given location, filled or unfilled.
        \item Display coloured text at any position.
        \item Change the colour of the background/canvas.
        \item Clear the screen of all drawings instantly.
      \end{itemize}
      An extra `bonus' feature would be to save the image drawn to file.
      Graphics were achieved in previous versions using Python's Turtle library. This was very effective
      for small simple simulations. However, when displaying flares on stares or when drawing lots of particles
      the performance drops rapidly. Ideally, an additional API such as OpenGL should be used.

      The picture to be drawn by the API will be provided by the camera. Like the simulator, the camera should
      only have to do the bare mininum work before handing over to the graphics API. The only information provided
      to the api from the camera will be:
      \begin{itemize}
        \item Type of shape.
        \item Screen position of shape.
        \item Shape parameters. (If required)
        \item Shape fill options. (If required)
      \end{itemize}
      Potentially, the brightness or light intensity of the object will also be given.

  \chapter{Simulation Algorithms}
    \section{Matrix Mechanics}
      Generally using a static n-dimensional array as a matrix will be
      faster than using a dynamic n-dimensional list. This is especially
      the case with Numpy's arrays compared to the native python lists.

      To make calculations we require the following arrays to be known at all times:
      \begin{itemize}
        \item $\bb P_N \in \mathbb{R}^{N \times 3} \to$ Array of position vectors.

        \item $\bb R \in \mathbb{R^N} \to $ Array of radii of particles.

        \item $\bb M \in \mathbb{R^N} \to $ Array of masses of particles.
      \end{itemize}
      Where $N$ is the number of particles. For a calculation such as standard
      gravitational attraction, we would perform the following particle for each particle.
      Here the individual particle we are simulating for is as position $\bb P$,
      with mass $M$ and radius $R$.
      \begin{align*}
        \intertext{Find the distance to each other particle:}
        \bb D_N &= \bb P_N - \{\bb P \} \\
        \intertext{(The braces indicate the subtraction is for each element in the matrix $\bb P_N$)}
        \intertext{The absolute distances are given by:}
        \bb D &= |\bb D_N| = \{|\bb x| \forall \bb x \in \bb D_N \}
        \intertext{For the force:}
        \bb F_N &=  G M \odot \bb M \odot (\bb D_N) \oslash (\bb D^3) \\
        \intertext{Where $\odot, \oslash$ are element wise multiply and divide respectively.
                  Also note that cubic power at the right is element wise aswell.}
        \intertext{Now we sum the forces to get the net force:}
        \bb F &= \sum_{n = 0}^{N} \bb F_{N_n} \\
        \intertext{The acceleration is simply derived from $\bb F = m \bb a$, however it
                  would be fine to simply remove the mass of the particle at the start
                  in this case to directly calculate acceleration instead of force.}
        \intertext{We also want to check for collisions. Lets consider a vector $\bb A$
                  containing the `altitude' of each particle, ie the surface-surface distance.}
        \bb A &= \bb D - \bb R - \{R\} \\
        \bb B_\mathrm{collision} &= \{\bb A < 0\}
        \intertext{Where $\bb B$ is a boolean vector where a value of True indicates that
                    our current particle is colliding with the particle at that index.}
      \end{align*}











\end{document}
